COVID-19 Policy-Compliance in Europe

### 2020-12-26 Draft.Rmd
Replaced automatically created regression tables with manually created markdown table. Added PDF with embedded fonts.

It yields the following result. Except for Date, non of the continuous variables Publicity and Cases/100.000/14 Days are statistically significant. Date is statistically significant at the 0.05 level. It's coefficient shows that when holding Publicity, Cases and Country constant, Compliance decreases by 0.07 points each day.

@@

```{r Regression Table, echo=FALSE, fig.align='center', results="asis"}

# Fit the model
lm1 <- lm(data = data, ComplianceIndex ~ Date + Publicity + notification_rate_per_100000_population_14.days + CountryName)

# Generate regression table
texreg::texreg(lm1, single.row = TRUE, 
#                         custom.header = list("Model Results" = 1),
#                         model.names = c("My name 1", "My name 2"), 
                         custom.coef.names = c("Intercept", "Date", "Publicity",
                                               "Cases/100,000/14 Days",
                                               "Belgium",
                                               "Denmark",
                                               "France",
                                               "Germany",
                                               "Ireland",
                                               "Italy",
                                               "Netherlands",
                                               "Poland",
                                               "Spain",
                                               "Sweden",
                                               "United Kingdom"),
                          bold = 0.05,
                          center = TRUE,
                          caption = "Regression Table"
                         )
```

@@

However, when plotting the fitted values from the non-hierarchical linear model by country, we can observe different discrete levels of compliance by date, e.g. with Austria showing much higher levels of compliance than Germany and Sweden.

```{r Linear model predictions by country, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
### Visualising different predicted intercepts and different slopes
ggplot(data = augment(lm(data = data, 
                         ComplianceIndex ~ Date + CountryName + Publicity + notification_rate_per_100000_population_14.days)),
       aes(x = Date, y = ComplianceIndex, color = CountryName))  +
  geom_smooth(aes(y = .fitted), method = "lm", se = FALSE) +
  xlab("Month") +
  scale_x_date(breaks = date_breaks("3 months"),
               labels = date_format("%b\n%Y")) +
  scale_y_continuous(name = "Policy Compliance Index") +
  theme(
    axis.title.y = element_text(color = "black", size=12, vjust = 3),
    axis.title.x = element_text(size=12, vjust = 0.5),
    text=element_text(family="CMU Serif")
    ) +
  ggtitle("Model Predictions for Policy Compliance in Europe Across Time")
```

This suggests creating a hierarchical model with a random intercept for country in order to improve model fit. We can also see that the slopes do not differ substantially, so that for the sake of simplicity we can use only a random intercept in our model.

\begin{equation}
{Compliance Index}_{ij} = {\beta}_0 + {\beta}_1 * {Date}_{ij} + {\beta}_2 * {Publicity}_{ij} + {\beta}_3 * {Cases/100,000/14 days}_{ij} + {Country}_j + {\epsilon}_{ij}
\end{equation}

```{r Regression Table 2, echo=FALSE, fig.align='center', results="asis"}

# Fit the model
lmer_out <- lmer(data = data, ComplianceIndex ~ Date + Publicity + notification_rate_per_100000_population_14.days +  (1 | CountryName))

# Generate regression table
texreg::texreg(lmer_out, single.row = TRUE, 
                         #                         custom.header = list("Model Results" = 1),
                         #                         model.names = c("My name 1", "My name 2"), 
                         custom.coef.names = c("Intercept", "Date", "Publicity",
                                               "Cases/100,000/14 Days"),
                                               #"Belgium",
                                               #"Denmark",
                                               #"France",
                                               #"Germany",
                                               #"Ireland",
                                               #"Italy",
                                               #"Netherlands",
                                               #"Poland",
                                               #"Spain",
                                               #"Sweden",
                                               #"United Kingdom"),
                         bold = 0.05,
                         center = TRUE,
                         caption = "Regression Table 2"
)

>>>>

It yields the following result. Except for Date, non of the continuous variables Publicity and Cases/100.000/14 Days are statistically significant. Date is statistically significant at the 0.05 level. It's coefficient shows that when holding Publicity, Cases and Country constant, Compliance decreases on average by 0.07 points each day.

| Variable              | Coefficient | Std. Error | P-Value |
| --------------------- | -----------: | ----------: | :-------: |
| Date                  | −0.07       | 0.01     | ∗∗∗     |
| Publicity             | −0.01       | 0.02     |         |
| Cases/100,000/14 Days | 0.00        | 0.00     |         |
| Belgium               | −3.21       | 2.12     |         |
| Denmark               | -8.29       | 2.12       | ∗∗∗     |
| France                | -7.15       | 2.11     | ∗∗∗     |
| Germany               | −11.06      | 2.12     | ∗∗∗     |
| Ireland               | −4.07       | 2.13     |         |
| Italy                 | −8.12       | 2.17     | ∗∗∗     |
| Netherlands           | −2.97       | 2.14     |         |
| Poland                | −5.67       | 2.17     | ∗∗      |
| Spain                 | −4.79       | 2.24     | ∗       |
| Sweden                | −9.53       | 2.11     | ∗∗∗     |
| United Kingdom        | −3.25       | 2.13     |         |
R\textsuperscript{2} = 0.31;
Adj. R\textsuperscript{2} = 0.29;
Num. obs. = 520;
∗∗∗ p < 0.001; ∗∗ p < 0.01; ∗ p < 0.05

@@

@@

As we can see in the regression output of our non-hierarchical model, the fixed effects estimated for the different expressions of our country variable show substantial differences, with Germany showing more than 10 points less policy-compliance than our reference country Austria.

Additionally, the following plot shows slightly different slopes of the countries.

```{r Non-hierarchical different intercepts and slopes, message=FALSE, warning=FALSE}
ggplot(data = augment(lm(data = data, 
                         ComplianceIndex ~ Date + CountryName + Publicity + notification_rate_per_100000_population_14.days)),
       aes(x = Date, y = ComplianceIndex, color = CountryName))  +
  geom_smooth(aes(y = .fitted), method = "lm", se = FALSE) +
  xlab("Month") +
  scale_x_date(breaks = date_breaks("3 months"),
               labels = date_format("%b\n%Y")) +
  scale_y_continuous(name = "Policy Compliance Index") +
  theme(
    axis.title.y = element_text(color = "black", size=12, vjust = 3),
    axis.title.x = element_text(size=12, vjust = 0.5),
    text=element_text(family="CMU Serif"),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey", linetype = "dotted")
  ) +
  ggtitle("Model Predictions for Policy Compliance in Europe Across Time")
```

Thus, in order to better assess the trends of policy-compliance in Europe over time, we can create a hierarchical model, allowing both random intercepts for the subgroups of Country as our grouping variable and random slopes along the Date variable.

\begin{equation}
{Compliance Index}_{ij} = {\beta}_0 + ({\beta}_1 * {Date}_{ij}) + {\beta}_2 * {Publicity}_{ij} + {\beta}_3 * {Cases/100,000/14 days}_{ij} + {Country}_{0j} + {\epsilon}_{0ij}
\end{equation}
